---
title: "Portfolio Item 2"
format:
  revealjs:
    theme: serif
    slide-number: true
    footer: "Exemplary code"
---

# Example Code

I chose the interactive graph from Problem Set 2 to be my example code. This was a data visualization project assigned on week 3 of the semester where we replicated <u> [this graph by Silver Bulletin](https://www.natesilver.net/p/trump-approval-ratings-nate-silver-bulletin)  </u> to the best of our abilities. 

## Process

As part of **Problem Set 2** we had to replicate a graph, the process included:

-   Accessing the data
-   Experimenting with code to replicate the graph
-   Adding an interactive layer to the process

# Setup

:::: {.columns}

::: {.column width="30%"}
Loading Packages

```{r, echo=TRUE}

# Load Packages
library(tidyverse)
library(ggplot2)
library(ggtext)
library(scales)
library(patchwork)
library(plotly)
library(reshape2)
library(readxl) 
```

:::

::: {.column width="70%"}
Creating my signature theme

```{r, echo=TRUE}
my_theme<- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "#FBFAF8", color = NA),
    panel.background = element_rect(fill = "#FBFAF8", color = NA),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "#E8E3DC", linewidth = 0.4),
    axis.title = element_text(family = "sans", color = "#4A4A4A"),
    axis.text = element_text(color = "#6A6A6A"),
    axis.ticks = element_blank(),
    plot.title = element_text(face = "bold", size = 19, color = "#2F2F2F", hjust = 0.3),
    plot.subtitle = element_text(size = 12, color = "#6F6F6F", hjust = 0.25),
    plot.caption = element_text( size = 9, color = "#8A8A8A", hjust = -1),
    legend.position = "right",
    legend.title = element_text(face = "bold", family = "sans"),
    legend.background = element_blank(),
    legend.key = element_rect(fill = NA)
  )
```

:::

::::

# Non-interactive graph

## Code

```{r, echo=TRUE}
kSCt4 <- read_excel("kSCt4.xls")

approval <- kSCt4|>
  ggplot( aes( x= modeldate) ) +
  geom_ribbon(
    aes(ymin = disapprove_lo, ymax = disapprove_hi, fill = '#16877a'),
    alpha=0.35
  )+
  geom_ribbon(
    aes(ymin = approve_lo, ymax = approve_hi, fill = '#d96003' ),
    alpha=0.35
  )+
  geom_line(
    aes( y= disapprove_lo ),
    color= '#d96003',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= disapprove_hi ),
    color= '#d96003',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= approve_lo ),
    color= '#16877a',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= approve_hi),
    color= '#16877a',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_hline(
    yintercept = 50
  )+
  geom_line(
    aes( y= disapprove ),
    color= '#d96003',
    alpha = 1,
    size = 0.8
  )+
  geom_line(
    aes( y= approve ),
    color= '#16877a',
    alpha = 1,
    size = 0.8
  ) +
  geom_text(
    data = kSCt4 |>
      slice_tail(n = 1),
    aes(x = modeldate, y = disapprove, label = paste0("Disapprove\n", round(disapprove, 1), "%")),
    color = '#d96003',
    size = 3.5,
    hjust = -0.1,
    vjust = 0.8
  ) +
  geom_text(
    data = kSCt4 |>
      slice_tail(n = 1),
    aes(x = modeldate, y = approve, label = paste0("Approve\n", round(approve, 1), "%")),
    color = '#16877a',
    size = 3.5,
    hjust = -0.1,
    vjust = 0.8
  ) +
  labs(
    title= "Do Americans <span style = 'color: #16877a ;'>approve </span>or <span style = 'color: #d96003 ;'>dispprove</span> of
Donald Trump?"  ,
    subtitle = "An updating polling average of Donald Trump's approval rating and disapproval rating in his
second term, accounting for each poll's quality, recency, sample size, and partisan lean.",
    caption = "Note: 90% of polls are projected to fall within the shaded regions.
Updated February 17, 2026 at 11:04 PM . Get the data . Download image
") +
  scale_x_date(
    limits = c(as.Date("2025-01-20"), as.Date("2026-02-11")),
    breaks = seq(as.Date("2025-04-01"), as.Date("2026-01-01"), by = "3 months"),
    labels = c("Apr 2025", "Jul", "Oct", "Jan 2026"),
    expand = expansion( mult = c(0, .1))
  )+
  scale_y_continuous(
    limits = c(30,70),
    labels = function(x) paste0(x, ".0%"),
    expand = expansion( mult = c(0, .1))
  )+
  guides(fill = "none", color = "none")  +
  annotate("label",
           x = as.Date("2025-03-15"),
           y = 70,
           label = "Approval/disapproval",
           fill = "black",
           color = "white",
           label.size = 0,
           label.r = unit(0.3, "lines"),
           fontface = "bold") +
  annotate("label",
           x = as.Date("2025-06-15"),
           y = 70,
           label = "Net approval",
           fill = "white",
           color = "black",
           label.size = 0.5,
           label.r = unit(0.3, "lines"),
           fontface = "bold") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_markdown(size=25),
    plot.title.position = "plot",
    plot.subtitle = element_text(hjust = 0, size = 17, family = "sans"),
    plot.background = element_rect(fill = "#f2f2f2", color = NA),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0, face = "italic", family = "sans"))

```

## Result

```{r}
approval
```

# Interactive graph

## Code

```{r, echo=FALSE}

approval1 <- kSCt4|>
  ggplot( aes( x= modeldate) ) +
  geom_ribbon(
    aes(ymin = disapprove_lo, ymax = disapprove_hi, fill = '#16877a'),
    alpha=0.35
  )+
  geom_ribbon(
    aes(ymin = approve_lo, ymax = approve_hi, fill = '#d96003' ),
    alpha=0.35
  )+
  geom_line(
    aes( y= disapprove_lo ),
    color= '#d96003',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= disapprove_hi ),
    color= '#d96003',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= approve_lo ),
    color= '#16877a',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_line(
    aes( y= approve_hi),
    color= '#16877a',
    linetype = 'dotted',
    size = 0.8,
    alpha = 0.6
  )+
  geom_hline(
    yintercept = 50
  )+
  geom_line(
    aes( y= disapprove ),
    color= '#d96003',
    alpha = 1,
    size = 0.8
  )+
  geom_line(
    aes( y= approve ),
    color= '#16877a',
    alpha = 1,
    size = 0.8
  ) +
  scale_x_date(
    limits = c(as.Date("2025-01-20"), as.Date("2026-02-11")),
    breaks = seq(as.Date("2025-04-01"), as.Date("2026-01-01"), by = "3 months"),
    labels = c("Apr 2025", "Jul", "Oct", "Jan 2026"),
   expand = expansion(mult = c(0, .18))
  )+
  scale_y_continuous(
    limits = c(30,70),
    labels = function(x) paste0(x, ".0%"),
    expand = expansion( mult = c(0, .1))
  )+
  guides(fill = "none", color = "none")  +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_markdown(size=25),
    plot.title.position = "plot",
    plot.subtitle = element_text(hjust = 0, size = 17, family = "sans"),
    plot.background = element_rect(fill = "#f2f2f2", color = NA),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0, face = "italic", family = "sans"))

```


```{r,echo=TRUE, eval=TRUE, fig.width=20, fig.height=8, out.width="100%"}

final <- ggplotly(approval1) |>
 layout(
   paper_bgcolor = "#f2f2f2",
   plot_bgcolor  = "#f2f2f2",
  title= list(
    text = paste0(
     "Do Americans <span style='color:#16877a'>approve</span> or <span style='color:#d96003'>disapprove</span> of Donald Trump?",
     "<br>",
      "<sub>An updating polling average of Donald Trump's approval rating and disapproval rating in his<br>second term, accounting for each poll's quality, recency, sample size, and partisan lean.</sub>"
     ),
     x = 0.03,
    xanchor = "left",
    font = list(size = 28 ),
    y = 0.90
    ),
  margin = list(t = 130, b = 62, r = 0),
  autosize = TRUE,
      annotations = list(
      list(
        x = as.numeric(as.Date("2026-02-11")),
        y = last(kSCt4$disapprove),
        text = paste0("Disapprove<br>", round(last(kSCt4$disapprove), 1), "%"),
        xanchor = "left",
        yanchor = "middle",
        showarrow = FALSE,
        font = list(color = "#d96003", size = 12),
        xref = "x",
        yref = "y"
      ),
      list(
        x = as.numeric(as.Date("2026-02-11")),
        y = last(kSCt4$approve),
        text = paste0("Approve<br>", round(last(kSCt4$approve), 1), "%"),
        xanchor = "left",
        yanchor = "middle",
        showarrow = FALSE,
        font = list(color = "#16877a", size = 12),
        xref = "x",
        yref = "y"
      ),
      list(
      text = "Note: 90% of polls are projected to fall within the shaded regions.<br>Updated February 17, 2026 at 11:04 PM · Get the data · Download image",
      x = -0.06,
      y = -0.06,
      xref = "paper",
      yref = "paper",
      xanchor = "left",
      yanchor = "top",
      showarrow = FALSE,
      font = list(size = 18, color = "black", face = "italic", family = "sans"),
      align = "left"
    )
    )
  ) |>
  config(responsive = TRUE)
```

## Result

```{r}
final
```

# Thank you
